Algebraic Groups\vspace{-1ex}

### 2019-11-08 Algebraic Geometry/Algebraic-Geometry.tex
update

	Then we can choose $f_0,\dots,f_m\in k[x_0,\dots,x_n]$ that are homogeneous of degree $d$ and these coordinate functions give us a rational map (in fact, all the rational maps) from $\P n$ to $\P m$.
\end{rmk}
\begin{prob}
	Find a morphism $f:\P1\to\P2$ whose image is $Z(xy-z^2)$.
\end{prob}
I am thinking $x_0^2$, $x_1^2$, and $x_0x_1$.
\begin{prob}
	Suppsoe that $Q$ is a non-singular quadratic form in three variables, so $Z(Q)\subseteq \P2$. Then show that $Z(Q)\cong \P1$.
	\textit{Hint: Linear changes of coordinates are algebraic!}
\end{prob}
\end{document}

>>>>

	Then we can choose $f_0,\dots,f_m\in k[x_0,\dots,x_n]$ that are homogeneous of degree $d$ and these coordinate functions give us a rational map (in fact, all the rational maps) from $\P n$ to $\P m$.
\end{rmk}
\begin{prob}
	Find a morphism $f:\P1\to\P2$ whose image is $Z(xy-z^2)$.
\end{prob}
I am thinking $x_0^2$, $x_1^2$, and $x_0x_1$.
\begin{prob}
	Suppsoe that $Q$ is a non-singular quadratic form in three variables, so $Z(Q)\subseteq \P2$. Then show that $Z(Q)\cong \P1$.
	\textit{Hint: Linear changes of coordinates are algebraic!}
\end{prob}

\section{November 8th, 2019}
Last time we were thinking about conic curves in $\P2$, that is $Z(Q)$ where $Q$ is a nondegenerate quadratic form in $x_0,x_1,x_2$. One thing that people tried 
was to classify the ways that $C=Z(Q)$ intersects with copies of $\A2$. Then using these options, you can proceed. This leverages automorphisms of 
$\P2$ to reduce everything to a very finite set of cases.

Another idea was to do a linear change of coordinates that takes $C$ to $Z(xy-z^2)$. This amounts to doing some work on the matrix representing $Q$. Then answer it for a hyperbola.

For the other question, we were trying to find a good morphism $\P1\to\P2$ with image the hyperbola. The idea was the one I said. It can be seen relatively easily to be bijective.
\begin{rmk}
	Is a bijective morphism an isomorphism?! (No) The idea to consider is $R=k[x,y]/(y^2-x^3)\to k[t]$ where $k[t]$ is the normal closure of the former. Then if we are looking at the fiber over $0$, 
	we compute $R/(x,y)\cong k\to k[t]/(t^2,t^3)\cong k[t]/t^2$
\end{rmk}
\begin{rmk}
	Another idea that came up: if we are thinking about the geometry behind the homogenization/dehomogenization maps, notice that this tells us something about how 
	dehomogenizing is not a unique process!
\end{rmk}

\subsection{The Veronese Embedding}
Given $n$ and $d$, there is a (somewhat) canonical map 
\[\P n\to \P{\binom{n+d}{d}-1}\]
where the coordinate functions are the degree $d$ monomials in $x_0,\dots, x_n$. The image here is ``cut out by quadrics.'' These quadrics 
come from $MN=PQ$ for (not necessarily distinct) coordinate functions.

Recall the Segre embedding. When $n=m=1$, we get an embedding 
\[\P1\times \P1\to\P3\]
where 
\[(x_0:x_1),(y_0,y_1)\mapsto (x_0y_0,x_0y_1,x_1y_0,x_1y_1)=(z_0,z_1,z_2,z_3).\]
Here the relation is $z_0z_3=z_1z_2$.

Some problems:
\begin{prob}
	Show that the image of $\P1\times\P1$ in $\P3$ above is $Z(z_0z_3=z_1z_2)$.
\end{prob}
\begin{prob}
		Show that $\P1\times \P1\not\cong \P2$.
\end{prob}
\begin{prob}
	For any non-degenerate quadratic form $z_0,\dots,z_3$, do we have that $Z(Q)\cong \P1\times\P1$?
\end{prob}
\begin{prob}
	Are all the smooth quadratic hypersurfaces of a fixed dimension isomorphic? Are they all isomorphic to $(\P1)^d$?
\end{prob}
\begin{prob}
	What about cubic hypersurfaces?
\end{prob}

\end{document}

### 2019-11-08 Algebraic Groups/Algebraic-Groups.tex
update

>>>>

\section{November 8th, 2019}
We started by talking elections. Go Andrew Lewis. You can do it.

Recall that last time we had a result that said that we had an analog of the stabilizer $G_W\subseteq G$ 
representing the functor 
\[R\mapsto \{g\in G(R)|gW_R=W_R\}.\]
The idea was to take a basis for $W$ and complete it to one of $V$ and then to find a group that fixes $W$. We found 
that $G_W=V(\{a_{kj}\})$.
\begin{thm}[Chevalley]
	If $H\subseteq G$ is a subgroup of an algebraic group over $k$, then there exists a $G$ representation $V$ and a line $L\subseteq V$ such that $H=G_L$.
\end{thm}
\begin{prf}
	Let $\pi:\Gamma(G)\twoheadrightarrow\Gamma(H)$ be the map induced by $H\hookrightarrow G$. Let $q_1,\dots,q_n$ be
	generators of $\ker\pi=I\subseteq\Gamma(G)$, the regular representation. Take a finite dimensional representation $V$ containing $I$ in $\Gamma(G)$
	and pick a basis $e_1,\dots,e_s$ of $W=V\cap I$. Extend this basis to one of $V$ with the additional vectors denoted $f_1,\dots,f_t$.

	Now the image of the coaction gives us 
	\[e_i\mapsto\sum_k a_{ik}\otimes e_k+\sum_k b_{jk}\otimes f_k\]
	and let $I'=(b_{jk})$. We claim that $I=I'$. If this is true, then $H=G_W$ and $W\subseteq V$, so we set
	\[L=\bigwedge^{\dim W}W\subseteq \bigwedge^{\dim W}V\]
	and now $H= G_L$.

	To see the claim, consider that since $e_i\in I$, we get $\sum_ka_{ik}\otimes e_i\in \Gamma(G)\otimes I$.
	But then the comultiplication on $\Gamma(G)$ gives us that 
	\[m^\ast(I)\subseteq I\otimes \Gamma(G)+\Gamma(G)\otimes I\subseteq \Gamma(G)\otimes\Gamma(G)\]
	so we get that 
	\[\sum b_{ik}\otimes f_k\in I\otimes \Gamma(G)+\Gamma(G)\otimes I\]
	but since $f_{ij}\notin I$, this forces $b_{ik}\in I$. Thus $I'\subseteq I$.

	For the other containment I got behind.
\end{prf}
\subsection{Quotients}
Let $H\subseteq G$ be normal. Now the goal is to construct $G/H$ as an algebraic group. As a preview, we are going to use the last theorem 
to get a representation $V$ of $G$ such that $H=G_L$ such that $G\to \GL(V)$ descends to a representation $G\to \GL(V^H)$. Then $G/H$ will be the image of this map in $\GL(V^H)$.

\begin{defn}
	A subgroup $H\subseteq G$ of an algebraic group over $k$ is \textbf{normal} if for all $k$-schemes $S$, $H(S)\le G(S)$ is anormal subgroup.
\end{defn}
\begin{rmk}
	As an exercise one can show that if $G(k)\subseteq G$, then $H\subseteq G$ is normal if and only if for all $g\in G(k)$, $gHg^{-1}\subseteq H$.
\end{rmk}
\begin{lem}
	If $\pi:H\to G$ is a homomorphism of algebraic groups, then $\ker\pi\subseteq H$ is normal.
\end{lem}
To see the above, consider the kernel is the pullback:
\begin{center}
	\begin{tikzcd}
		S\ar[dr,dashed]\ar[drr,bend left]\ar[ddr,bend right] & &\\
		& \ker\pi\ar[r]\ar[d] & \Spec k\ar[d,"e_G"]\\
		& H\ar[r,"\pi"] & G
	\end{tikzcd}
\end{center}
And then the kernel $\ker \pi(S)=K(S)$ is normal in $H$. Think about this one some more.
\begin{defn}
	A homomorphism of algebraic groups $G\to Q$ is a \textbf{quotient map} if $G\to Q$ is faithfully flat.
\end{defn}
\begin{prop}
	A quotient map of linear algebraic groups $\pi:G\to Q$ satisfies the universal property: for all $f:G\to H$,
	\begin{center}
		\begin{tikzcd}
			K=\ker\pi\ar[r]\ar[rd,"0"] & G\ar[r,"\pi"]\ar[d,"f"] & Q\ar[dl,dashed,"!"]\\
			& H &
		\end{tikzcd}
	\end{center}
\end{prop}
